\documentclass[a4paper]{article}

\usepackage[margin=1.2in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{amsmath, amssymb}
\usepackage{amsthm}
\usepackage{braket}

\DeclareMathOperator{\R}{\mathbb{R}}
\DeclareMathOperator{\C}{\mathbb{C}}
\DeclareMathOperator{\N}{\mathbb{N}}
\DeclareMathOperator{\Z}{\mathbb{Z}}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\GF}{GF}
\DeclareMathOperator{\GR}{GR}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corrollary}
\newtheorem{example}{Example}

\title{Some notes}
\author{Ernesto Camacho Ramírez}
\begin{document}
  \maketitle
  
  \section{Investigations on discrete Wigner functions of
  multi-qubit systems}

  \subsection{K. Srinivasan}

  Uses Gibbons and Wootters method for constructing discrete
  Wigner functions. This means they assign mutually unbiased
  bases to straight line striations of the phase space. They
  acknowledge that by doing this, there is no unique Wigner
  function associated to any density matrix. 

  A Hilbert space of dimension $N$ is associated with a
  discrete phase space of $N \times N$ cardinality. The
  horizontal and vertical axes of the discrete phase space
  are asociated with two non-commuting observables. The
  points are ordered pairs of elements of a Galois field of
  size $N$. Given the equation for a line $aq + bp = c$, if
  we fix the values of $a$ and $b$ and vary $c$ over the
  field, then a set of $N$ parallel lines called a striation
  is generated. Each striation contains exactly one ray, and
  for fixed values of $x$ and $y$, varying  $s$ from the
  Galois field, then the set of $N$ points $(sa,sb)$ gives a
  ray.

  For prime power dimensions there are $N+1$ striations in
  the discrete phase space and $N+1$ MUBs. In order to give
  lines and striations a physical meaning, Wootters
  associates the lines of the phase space with pure states:
  if $\lambda_j^i$ is the $i$-th line of the $j$-th
  striation, then $Q(\lambda_j^i) =
  \ket{\lambda_j^i}\bra{\lambda_j^i}$ is the rank one
  projector associated with the line.

  This guy mentions that, just like the continuous case, one
  can define a set of $N^2$ translations $\mathcal
  T_{(q,p)}$ in discrete phase space, which shifts each
  point in a line by the amount $(q,p)$. For every
  striation, there exists $N-1$ translation operators which
  leave the lines in the striation invariant and those
  translation operators are formed by the points of the ray
  of the corresponding striation. So in order to achieve
  translational covariance, Wootters idea is to associate to
  each phase space translation a corresponding translation
  (displacement) operator, such that $T_\alpha T_\beta =
  \eta T_{\alpha+\beta}$ and such that we can decompose the
  translation operator as a tensor product of
  \textit{basic} translation operators that act upon the
  individual subystems of the composite Hilbert space:
  \[
    T_{(q,p)}
    = T_{(q_1,p_1)} \otimes T_{(q_2,p_2)} \otimes \cdots
    \otimes T_{(q_n,p_n)}.
  \] 

  For a prime dimension system, the generalized Pauli
  matrices $X$ and $Z$ can be associated to the unit
  horizontal and vertical translations $\mathcal T_{(1,0)}$ 
  and $\mathcal T_{(0,1)}$. Such operators obey the
  commutation relation $ZX = \eta XZ$. And so the unitary
  operators are then defined up to a phase factor as
  \[
    T_{(q,p)}
    = X^{q_1} Z^{p_1} \otimes \cdots \otimes X^{q_n}
    Z^{p_n}.
  \] 

  The condition of translational covariance is ensured by
  satisfying the equation
  \[
    Q(\mathcal T_\alpha \lambda)
    = T_\alpha Q(\lambda) T_\alpha^{*}.
  \] 
  Which in turn requires the translation operators
  corresponding to a striation (that is those which leave
  the parallel lines of the striation invariant), commute.

  Then in order to construct the Wigner function, we start
  with its fundamental property, that is the sum (integral)
  along a particular line $\lambda$ is equal to the
  expectation value of the rank one projector $Q(\lambda)$.
  The sum of the Wigner function $W$ along this line is
  equal to the probability $p(\lambda) = \Tr(Q(\lambda)
  \rho) = \sum_{\alpha \in \lambda}^{} W_\alpha$, so we can
  reconstruct the Wigner function by the $N+1$ projective
  measurements. Note that for any given point $\alpha$ in
  the discrete phase space, there are $N+1$ lines which pass
  through it. Then, the sum of Wigner elements $S(\alpha)$ 
  along the $N+1$ lines wich contains the point is given by
  \[
    S(\alpha)
    = \sum_{\lambda \ni \alpha}^{} \sum_{\beta \in
    \lambda}^{} W_\beta
    = NW_\alpha + \sum_{\gamma}^{} W_\gamma.
  \] 
  Then
  \[
    W_\alpha = \frac{1}{N} \left( \sum_{\lambda \ni
    \alpha}^{} \sum_{\beta \in \lambda}^{} W_\beta -
  \sum_{\gamma}^{} W_\gamma \right), 
  \] 
  where $\sum_{\beta \in \lambda}^{} W_\beta$ is the
  probability to measure the systems along the projector
  $Q(\lambda)$. And so
  \[
    W_\alpha
    = \frac{1}{N} \Tr(A_\alpha \rho),
  \] 
  where the $A_\alpha$ are the so called phase-point
  operators, given by
  \[
    A_\alpha = \sum_{\lambda \ni \alpha}^{} Q(\lambda) - I.
  \] 

  The author then goes about providing the phase-point
  operator properties along with Wigner function properties
  (note the Phd thesis does not even prove them). Then he
  talks about a labeling scheme for lines and striations in
  the particular case of multiqubit Wigner functions.
  Identifying the origin with the element $(0,0)$, there
  exists exactly one line in each striation passing through
  the origin called a ray. The rays are formed by the points
  $(sx,sy)$ for fixed $(x,y)$. In this case the rays of the
  horizontal and vertical striations are given by the points
  $(1,0)$ and $(0,1)$. The other rays are generated by
  $(1,\omega), (1,\omega^2), \ldots,(1,\omega^{N-2})$. And
  so he associates a \textit{striation number} $j$ with a
  fixed point and the ray $\lambda_j^0$. Given the ray of
  the vertical striation $\lambda_1^0$, the other lines of
  the striation are obtained using the set of $N-1$ 
  horizontal translation operators, $\mathcal T_{(s,0)}$,
  such that $\lambda_1^i = \mathcal T_{(i,0)}\lambda_1^0$.
  The author \textit{chooses} the translations $\mathcal
  T_{(0,s)}$ for obtaining the other lines by acting on the
  ray of the rest of the striations.

  The author continues by given a concrete example: a single
  qubit system. As we know, the horizontal and the vertical
  axes are associated with the Pauli matrices $\sigma_x$ and
  $\sigma_z$, which are non-commuting. There are $6$ lines
  in the phase space, which can be grouped into 3 sets of
  parallel lines. The corresponding translation operators
  are associated to the points $(0,s)$, $(s,0)$ and $(s,s)$.
  The translation operators are traceless and mutually
  orthogonal, and their eigenvectors forms the MUBs. The
  asociation of the rank one operators and the lines is not
  unique, in fact there are $2^{2+1} = 8$ possible ways to
  do the assignment. After a choice of quantum net, we can
  build the phase-point operators and then calculate the
  Wigner function.

  For the case of the two qubit system, for each striation,
  there exists a set of $N-1 = 3$ translations which leave
  the lines of the striation invariant. The translations
  associated to the five striations are $\mathcal T_{(0,s)},
  \mathcal T_{(s,0)}, \mathcal T_{(s,s)}, \mathcal
  T_{(s,s\omega}$ and $\mathcal T_{(s,s\omega^2)}$. There
  are 15 unitary translation operators along with the
  identity. The oeprators can be equally partitioned into
  $5$ sets and each set contains $3$ operators. The common
  eigenstates of the $3$ traceless, mutually orthogonal,
  commuting operators are associated with the lines of the
  striation. In this case there are $4^{4+1} = 1024$ quantum
  nets.

  After this, the author talks about quantum nets and
  equivalence classes as studied by Gibbons and Wootters.
  Then he starts talking about the advantages and
  limitations of the Gibbons construction. The main
  disadvantage is that the Wigner function as constructed by
  Wootters is not unique, as there are $N^{N+1}$ versions of
  it. As such, the non-classicality of the reconstructed
  state is not obvious. Interestingly he shows the results
  of an experiment of the tomographic reconstruction of the
  experimental data.

  \section{Geometrical Representations of Finite Dimensional
  Quantum Systems}

  \subsection{Daniël Marjenburgh}

  The authors plan to introduce a discrete Wigner function
  is the same as before: to preserve as many of the
  continuous version's properties as possible. He starts
  from first principles and the phase point operators, using
  the traditional definition of the Wigner function as the
  trace of product of the density matrix and the displaced
  parity operators. The properties that are imposed are that
  of translation covariance and the tomographic property.

  The author decides to use the $N \times N$ grid of points
  as the natural phase space, noting that there is no
  redundency in the case of quantum tomagraphy (unlike the
  DWF that use a $2N \times 2N$ phase space).

  This guy gives a very intuitive explaination for the
  natural analog of the continuous, the conjugate bases and
  the property of mutual unbiasedness, along with the
  Fourier transform.

  Interesting conclusions on the even and odd cases, noting
  that for $N$ odd, there is a unique expression for the
  phase point operators such the WF satisfies the
  tomographic property and translation covariance.

  \section{The interplay between quantum contextuality and
    Wigner negativity}
  
  \subsection{Pierre-Emmanuel Emeriau}

  The author's thesis focuses on two nonclassical
  behaviours: quantum contextuality and Wigner negativity.
  The is a notion superseding nolocality that can be
  exhibited by quantum systems. In discrete settings,
  contextuality has been shown to be necessary and
  sufficient for advantages in some cases. This seems to be
  a continous Wigner function focused thesis. 

  \section{Hudson's theorem for finite-dimensional quantum
  systems}

  \subsection{D. Gross}

  Gross shows that for a system of odd dimension, only the
  stabilizer states posses a non-negative Wigner function.
  The give an axiomatic characterization which completely
  fixes the definition of the Wigner function.

  Gross notes that partly triggered by the advent of quantum
  information theory, considerable work has been undertaken
  to explore Wigner functions for finite-dimensional quantum
  systems. Finally someone that agrees with me, Gross notes
  that there seem to be to clear distinct approaches that
  can be identified in the leterature. The first tries to
  cast the definition of the Wigner function into a form
  that can be interpreted for both continuous variable and
  discrete systems, these are the works of Vourdas (2004),
  Paz (2002), Klimov and Gross. The second approach,
  introduced by Gibbons, Hoffman and Wootters, is based on
  the properties of the continuous Wigner function.

  Gross actually proposes a definition for odd dimensions
  that highly resembles the continuous version and
  supposedly is most sensible judged in both approaches. He
  defines it as:
  \[
    W_\psi(p,q)
    = d^{-1} \sum_{\xi \in \Z_d}^{} e^{-\frac{2\pi}{d} i \xi p}
    \bar{\psi}(q-2^{-1}\xi) \psi(q+2^{-1}\xi).
  \] 
  Such definition is actually the disrete symplectic Fourier
  transform of the discrete characteristic function (like
  Vourdas' approach) and he shows that it is the unique
  choice that mimics certain desirable properties.

  He proceeds by talking about stabilizer states, which are
  joint eigenvectors of certain sets of elements of the
  qubit Pauli group. There is of course, stabilizer states
  for higher-dimensional quantum systems. Although they
  display complex features such as multi-particle
  entanglement, such states allow for an efficient classical
  description. The implications of the Gottesman-Knill
  theorem are quite important. He then proposes the
  following theorem:

  \begin{theorem}
    Let $d$ be odd and $\psi \in L^2(\Z_d^n)$ be a state
    vector. The Wigner function of $\psi$ is non-negative if
    and only if $\psi$ is a stabilizer state. Furthermore,
    since $\psi(q) \neq 0$ for all $q$, a vector $\psi$ is a
    stabilizer state if and only if it is of the form
    \[
      \psi(q) \propto e^{\frac{2\pi}{d}i (q\theta q + x q)},
    \] 
    where $q,x \in \Z_d^{n}$ and $\theta$ is a symmetric
    matrix with entries in $\Z_d$.
  \end{theorem}

  Gross' theorem shows that if the \textit{right}
  definitions are employed, the continuous and discrete case
  behave very similarly. Also one can further the idea that
  stabilizer states are like Gaussian states.

  In order to vinculate Galvao's research on the
  non-negativity of the DWF, Gross comments on Wootters
  approach. The axioms such definition should meet are:
  \begin{itemize}
    \item $W$ is a linear mapping sending operators to
      functions on a $d \times d$ lattice, called the phase
      space.
    \item The Wigner function is covariant under the action
      of the Weyl operators.
    \item There exists a function $Q(\lambda)$ that assigns
      a pure quantum state to every line $\lambda$ in phase
      space. If $\psi$ is a state vector, then the sum of
      its Wigner function along $\lambda$ must be equal to
      the overlap $|\braket{Q(\lambda)|\psi}|^2$.
  \end{itemize}

  Gross notes that for a $d$-dimensional Hilbert space,
  there exist $d^{d+1}$ distinct generalized Wigner
  functions, and that such approach is only described for $d
  = p^{n}$ where $p$ is a prime number, beacuase the of the
  necessary finite geometry given by finite fields.

  Next up is the notion of the stabilizer states. Consider a
  composite system, build of $n$ $d$-level particles.
  Considering the system as a single $d^{n}$-dimensional
  space or as a composite system gives distints notions of
  stabilizer states. Galvao's work concerns quantum states
  in prime power dimensions that are non-negative with
  respect to all possible definitions of generalized Wigner
  functions. Such states are shown to be mixtures of
  single-particle stabilizer states. If the Wigner function
  of a quantm computer is positive at all times, then it
  operates only with stabilizer states. For the case of
  non-qubit pure states, Gross' theorem implies Galvao's
  results, and favorably, there is only one definition to
  study. Although Gross' theorem does not address qubits or
  mixed states.

  The next section gives a review of the Weyl
  representation. Start with an odd $d$-dimensional quantum
  system. Choose the computational basis labeled by the
  elements of the ring $\Z_d$. The main objects in the phase
  space formulation are the Weyl operators (generalized
  Pauli operators). Define (the character) $\chi(q) =
  e^{\frac{2\pi}{d}iq}$ and define the shift and boost
  operatros:
  \[
    \hat x(q) \ket x = \ket{x+q}
    \quad
    \hat z(p) \ket x = \chi(px) \ket x.
  \] 
  The Weyl operators are then defined as (just like Vourdas)
  \[
    w(p,q) = \chi(-2^{-1}pq) \hat z(p) \hat x(q),
  \] 
  for $p,q,t \in Q = \Z_d$. Gross notes the choice of phase
  factors ensures that the symplectic inner product appears
  in the composition law, thus making the connection between
  the Weyl operators and symplectic geometry manifest. Other
  definitions (like no phase factor) carry the same
  dependence in a less obvious manner. The set of Weyl
  operators is closed under multiplication, up to phase
  factors. Writing  $w(v) = w(v_p,v_q)$ for elements $v =
  (v_p,v_q) \in \Z_d^2$, then $V := \Z_d \times \Z_d$ with
  the symplectic inner product is called the phase space.

  Such construction can be generalized to multi particle
  systems. Consider $\Z_d^{n}$ where the product is given by
  the inner product. The Weyl operators are now given by the
  tensor products
  \[
    w(p,q)
    = w(p_1,\ldots,p_n,q_1,\ldots,q_n)
    = w(p_1,q_1) \otimes \cdots \otimes w(p_n,q_n).
  \] 

  Finally he gives some remarks that were left unmentioned.
  A state vector $\ket \psi$ can be identified with a
  complex function on $\Z_d$ using the inner product
  $\psi(q) = \braket{q|\psi}$.

  Next Gross introduces the Clifford group, as the set of
  unitary operators such that
  \[
    Uw(v)U^{*} = c(v)w(S(v))
  \] 
  for some maps $c: V \to \C$ and $S : V \to V$. For prime
  values of $d$, $\Z_d$ has a finite field structure, and
  $\Z_d^{n}$ is a finite vector space. He goes into some
  detail about the Clifford group and affine
  transformations, the metaplectic group.

  Next he sets up the Fourier transforms on the space
  $\Z_d^n$ for a complex function $f: Q \to \C$. He defines
  it as
  \[
    \mathcal Ff(p)
    = (d^{n})^{-1 / 2} \sum_{q \in \Z_d^{n}}^{}
    \hat\chi(pq)f(q).
  \] 
  Then he defines a more robust version using the characters
  $\zeta$.

  Finally he proceeds to define and give properties of the
  Wigner function. The Weyl operators $\{w(p,q)\}$ form an
  orthonormal basis in the space of operators on $\mathcal
  H$ with respect to the trace scalar product. Just like in
  the continuous case, the \textit{characteristic function}
  of the operator $\rho$ is given by its expansion
  coefficients with respect to the Weyl basis:
  \[
    \Xi_\rho(\xi,x)
    = d^{-n} \Tr(w(\xi,x)^{*}\rho).
  \] 
  And now he completes the analogy with the continuous case,
  as he has defined the symplectic Fourier transform and the
  characteristic function. And so for $d$ odd, $Q =
  \Z_d^{n}$, let $V, \mathcal H$ be as usual and let $\rho$ 
  be a quantum state. Then the Wigner function $W_\rho$ is
  the symplectic Fourier transformation of the
  characteristic function $\Xi_\rho$. An explicit
  calculation gives the usual definition:
  \[
    W_\rho(a)
    = d^{-n} \Tr(A(a)\rho),
  \] 
  where the $A(a)$ supposedly coincides with Wootters phase
  space point operators. He then states the usual properties
  of the Wigner function: the phase point operators have
  unit trace and form an orthonormal basis in the space of
  Hermitian operators on $\mathcal H$. And so the Wigner
  function of a Hermitian operator is real and we have the
  overlap and normalization properties. Constructed in this
  way, the Wigner function for a pure state is given by the
  equation at the beginning, which looks like Wigner
  original. The marginal properties hold as well with
  respect to the conjugate variables. He also notes that the
  phase point operators factor as a tensor product
  \[
    A(p_1,\ldots,p_n,q_1,\ldots,q_n)
    = \bigotimes_i^n A^{(i)}(p_i,q_i)
  \] 
  and hence so does the Wigner function. The origin point
  operator is the parity operator and he even gives the
  Moyal product.

  Important for his theorem is the proof the Wigner function
  defined in his manner is Clifford covariant. Also, Gross
  mentions that his version coincides with Vourdas and some
  other ones, plus Klimovs version for a specific example.
  Also, his construction satisfies the axioms put forth by
  Wootters and Gibbons, noting that Wootters mentions that
  some of the generalized Wigner functions are special in
  some sense of symmetry. In his language, the symmetry is
  showed by the Clifford covariance. Importantly he notes
  that if we require a definition of a Wigner function to be
  Clifford covariant, then this condition practically picks
  out his construction, giving a sense a uniqueness.

  Gross goes back to work with the stabilizer states, first
  noting that using the composition law of the Heisenberg
  group, two Weyl operators $w(v_1),w(v_2)$ commute if and
  only iff $[v_1,v_2] = 0$. And then considering the image
  of an entire subspace $M$ under the Weyl representation
  $w$. The set
  \[
    w(M) = \{w(m) | m \in M\}
  \] 
  consists of mutually commuting operators if and only if
  the symplectic form vanishes on $M$:
  \[
    [m_1,m_2] = 0,
    \quad m_i \in M.
  \] 
  Such spaces are called \textit{isotropic}. If $M$ is
  isotropic, then the operators $w(M)$ can be simultaneously
  diagonalized. If $M$ is maximally isotropic subspace of
  $V$ and $v \in V$, then up to a global phase, there is a
  unique state vector $\ket{M,v}$ that fulfills the
  eigenvalue equations
  \[
    \chi([v,m])w(m)\ket{M,v} = \ket{M,v}
  \] 
  for all $m \in M$. Such states are called stabilizer
  states.

  I think the important thing to note here is that for every
  odd prime dimension, the Wootters discrete Wigner function
  is uniquely determined by Clifford covariance. And that no
  Clifford covariant operator basis or discrete Wigner
  function exists in any even prime power dimension.

  \section{Gross' thesis}

  Gross' approach is the same as Vourdas. First he discusses
  the notions of the Heisenberg group, the characteristic
  function and Wigner function. Then he contributes to an
  open problem in the theory of stabilizer codes. His way of
  doing it is to center around the connection to the
  continuous way.

  He starts by defining characters. If $G$ is a finite
  abelian group, a character $\chi$ of $G$ is a homomorphism
  from $G$ into the circle group $S^{1}$, that is the set of
  complex numbers of modulus one. The set of all characters
  becomes a group of its own, called $G$ 's dual group $\hat
  G$. Finite abelian groups are isomorphic to their
  respective dual groups but there is no canonic way of
  identifying  $G$ with $\hat G$.

  As a fundamental example he defines the characters on
  finite fields. For prime dimension we have
  \[
    a \mapsto \chi_a(\cdot) := \omega^{a}.
  \] 
  For a finite extension of a prime field $\mathbb F_p$ we
  have
  \[
    b \mapsto \chi_b(\cdot) := \chi_{\mathbb F_p}(\Tr(b
    \cdot)).
  \] 

  Next he introduces the notion of a finite symplectic
  geometry, which reduces to the theory of finite vector
  spaces with a symplectic form. A finite vector space $V$ 
  is called symplectic if it posseses a non-degenerate
  bilinear form $[\cdot,\cdot]$ which is anti-symmetric
  \[
    [v,w] = -[w,v].
  \] 
  Should $V$ be defined over a field of characteristic two,
  then we additionally demand the form fulfills
  \[
    [v,v] = 0.
  \] 
  Given a subspace $M$ of $V$, the symplectic complement
  $M^{\perp}$ is the set of all $v \in V$ such that $[v,m] =
  0$ for all $m \in M$. $M$ is \textit{isotropic} if the
  form vanishes on $M$. Of note is that fact that if $M$ is
  isotropic then $M \subset M^{\perp}$. The maximum
  dimension of an isotropic space is $\dim V / 2$. A space
  that reaches this limit is maximal isotropic. Gross then
  subsequently refers to symplectic vector spaces as phase
  spaces.

  Next he defines the Fourier transform and the symplectic
  Fourier transform for abelian groups. In particular, he
  writes the Fourier transform of an $f \in L^{1}(\mathbb
  F)$ as a function on $\mathbb F$ itself by choosing a
  faithful character $\chi$ of $\mathbb F$:
  \[
    \hat f(a)
    = \hat f(\chi_a)
    = \frac{1}{\sqrt{d}} \sum_{b \in \mathbb F}^{}
    \braket{\chi|ab}^{*}f(b).
  \] 
  By choosing a character $\chi$ of $\mathbb F$, then for a
  symplectic vector space $V$ over the finite field, the
  character induces an isomorphism from $V$ to $\hat V$ by
  \[
    \chi_a(\cdot) = \chi([a,\cdot]).
  \] 
  He then defines the symplectic Fourier transform. 

  In the next section, Gross starts by saying that if a
  quantum system posses a symmetry, then there should exist
  a unitary, irreducible, possibly projective representation
  of the symmetry group. He starts with a field $\mathbb F$ 
  that is not of characteristic two and defines the
  Heisenberg group $H(\mathbb F)$ by its composition law
  using the standard symplectic form on the vector space
  $\mathbb F^2$. The Heisenberg group enters the quantum
  scene through the Weyl representation which maps the group
  to operators on the Hilbert space $\C^{d}$ where $d$ is
  the order of the finite field. The representation is
  constructed by fixing a character $\chi$ of $\mathbb F$ 
  and choosing a basis in $\mathcal H$. Next he defines what
  he calls the shift and clock operators
  \begin{align*}
    x(q) &: \ket{\phi_k} \mapsto \ket{\phi_{k+q}} \\
    z(p) &: \ket{\phi_k} \mapsto \chi(pk) \ket{\phi_k}.
  \end{align*} 
  And so the Weyl representation is given by
  \[
    w(p,q,t) := \chi(t - 2^{-1}pq) z(p)x(q).
  \] 
  The image of the $w$ is the set of Weyl operators. Two
  Weyl operatos commute if the symplectic inner product
  vanishes. 

  The definition of the Heisenberg group extends to finite
  vector spaces $\mathbb F^{n}$ in a natural way,
  \[
    \mathcal J_{\mathbb F^{2n}} = \bigotimes_{i=1}^{n}
    \mathcal J_{\mathbb F^2}.
  \] 
  Using this we can define the group $H^{n}(\mathbb F)$. The
  Weyl representation of $H^{n}(\mathbb F)$ is then defined
  as
  \[
    (p,q,t)
    \mapsto
    \chi(t)w(p_1,q_1) \otimes \cdots \otimes w(p_n,q_n)
  \] 
  where $p_i$ and $q_i$ are the components of $p$ and $q$ 
  with respect to the natural basis in $\mathbb F^{n}$.

  Now, for the qubit case, the fields that should be used
  are of characteristic two, and so we cannot define the
  Heisenberg group which requires the use of the symbol
  $2^{-1}$. However, Gross shows us that there is still a
  projective representation of $\mathbb F_2$ defined in an
  analogous way:
  \[
    w(p,q) := i^{-pq}z(p)x(q),
  \] 
  for $p,q \in \mathbb F_2$. The group generated by
  $\{w(p,q)\}_{p,q}$ is called the Pauli group, or the
  Heisenberg group for qubits even though it does not
  satisfy the composition law definition. He also mentions
  that the Heisenberg group for qubits is an extension of
  $\Z_4$ by $\mathbb F_2^{2n}$ while the non-binary
  Heisenberg group extends $\mathbb F_d$ by $\mathbb
  F_d^{2n}$.

  He proceeds as in the other article, to define the
  characteristic function of an operator, given by the
  coefficients of the expansion in terms of the Weyl
  operators.

  Next he defines the Wigner function of a hermitian
  operator $A$ to be the symplectic Fourier transform of its
  characteristic function:
  \[
    W := \hat{\mathcal F} \circ \Xi.
  \] 
  Considering the Hilbert space $\mathcal H = \C^{nd}$, $d =
  p^{r}$, $V = \mathbb F_d^{2n}$, the Wigner function for
  all $a \in V$ is given by
  \[
    \frac{1}{d^{n}}
    \Tr\left( 
      \frac{1}{d^{n}} \sum_{b \in V}^{}
      \chi([a,b])^{*}w(b)^{*} A
    \right) .
  \] 
  This leads us the definition of the phase space point
  operators
  \[
    A(a)
    = \frac{1}{d^{n}} \sum_{b \in V}^{}
    \chi([a,b])^{*}w(b)^{*}.
  \] 
  These phase space point operators are hermitian, they have
  unit trace and are an orthonormal basis with respect to
  the Hilbert-Schmidt inner product. These properties
  immediately gives us an interpretation of the Wigner
  function. Furthermore, the Wigner function constructed in
  this manner has symplectic covariance. In the case of the
  qubit, the Wigner function looses its covariance under
  Clifford operations.

  Next he handles the case of extension fields. He considers
  the one-dimensional Heisenberg group on an extension field
  $\mathbb F_{p^{r}}$. Defining the character as
  $\chi(\cdot) = \omega^{\Tr a}$, he defines the reduced
  Heisenberg group with its own composition law. The Weyl
  representation of $H(\mathbb F_{p^{r}})$ is defined on
  \[
    \mathcal H = \C^{p^{r}}.
  \] 
  Introducing a tensor structure on $\mathcal H$, he defines
  the shift and clock oeprators, where these composed of
  operators that act only on specific subsystems:
  \[
    w(p,q,t) 
    = \bigotimes_i w^{(i)}(p_i,q^{i}),
  \] 
  and hence the Weyl representation factors with respect to
  the tensor structure.

  As an intereseting note, he mentions a natural question.
  Given an $p^{r}$ dimensional Hilbert space, is it better
  to associate it with a two dimensional phase space over
  $\mathbb F_{p^{r}}$ or a with a $2r$-dimensional one over
  $\mathbb F_p$? The latter choice seems to be more natural
  since all relevent structures (subspaces, isotropic
  spaces, symplectic mappings) can be mapped from $\mathbb
  F_{p^{r}}^{2}$ to $\mathbb F_p^{2r}$ but not the other way
  around. However, certain constructions in quantum state
  tomography and in the theory of mutually unbiased bases
  rely on the geometry of a two dimensional vector space
  which is of course more present in $\mathbb F_{p^{r}}^2$.

  \section{Codes, quadratic forms and finite geometries by
  Kantor}

  Kantor mentions that important types of projective planes
  are described using more general types of binary
  operations. Hence they are led to introduce those
  operations important in their coding-theoretic context.

  \begin{definition}
    Consider a binary operation $*$ on $\GF(2^{m})$ related
    to field multiplication by the following conditions:
    \begin{itemize}
      \item $(x+y) * z = x*z + y*z$.
      \item $x*y=x*z \implies x = 0$ or $y = z$.
      \item $\tr(x(x*y)) = \tr(xy)$.
      \item $x * y = 0 \iff x = 0$ or $y = 0$.
    \end{itemize}
  \end{definition}
  Kantor mentions that a fundamental aspect of the subject
  matter in his paper is that the first and second item in
  the previous definition are familiar in the theory of
  projective planes. These amount to some distorted versions
  of fields. It is just such field-like algebras that arise
  in the coordinatization of projective planes. Here he
  gives a recognized example:

  \begin{example}
    Let $T_1$ denote the trace map from $\GF(2^{m})$ to some
    proper subfield $F \neq \Z_2$. Then
    \[
      x * s = xs^2 + T_1(x)s + T_1(xs)
    \] 
    satisfies the definition of the ``algebra''.
  \end{example}

  He then starts with finite orthogonal geometries. He
  starts out with a vector space $V = \Z_2^{2n} = X \oplus
  Y$ for subspaces $X$ and $Y$ identified with $\Z_2^{n}$.
  He equips $V$ with a quadratic form given by the usual dot
  product, obtaining an associated bilinear form. He then
  defines totally singular subspaces of $V$. By a counting
  argument he notices that $V$ has $(2^{n}-1)(2^{n-1}+1)$ 
  nonzero singular vectors. Each totally singular $n$-space
  consists of singular vectors and contains $2^{n}-1$ 
  nonzero ones. This suggests that there might be families
  $\Sigma$ of totally singular $n$-spaces that partition the
  set of all nonzero singular vectors. Such a family of
  $2^{n-1}+1$ subspaces is called an orthogonal spread. He
  notes that such a family cannot exist unless $n$ is even,
  and that there is always at least one such family when $n$ 
  is even.

  Finally the definition of Kantor's $O^{+}(2n,2)$ notation
  is given. An isometry of $V$ is a nonsingular linear
  transformation preserving the quadratic form. These form a
  group, the \textit{orthogonal group} $O^{+}(2n,2)$ of $V$.
  The ``$+$'' refers to the fact that  $V$ has totally
  singular $n$-spaces. This group is transitive on the
  ordered pairs of totally singular $n$-spaces having only
  $0$ in common: from the point of view of this orthogonal
  geometry of $V$, the pair $X,Y$ we started with is
  indistinguisable from any other such pair. He then goes on
  to prove a couple of lemmas about the relationship between
  skew-symmetric matrices and orthogonal spreads. From which
  he obtains a corllary that allows him to construct
  orthogonal spreads of $V$ from Kerdock sets $\mathcal K$.
  Furthermore, he shows that every orthogonal spread of $V$ 
  that contains both $X$ and $Y$ arises in such a way.
  
  Inequivalent choices of $\mathcal K$ can produce
  equivalent orthogonal spreads. Naturally there are
  different notions of equivalence between pairs of Kerdock
  codes, pairs of Kerdock sets and pairs of orthogonal
  spreads.

  Next he shows how to construct orthogonal spreads from
  other geometric objects. First he shows how to obtain
  \textit{symplectic spreads} from orthogongal spreads. Then
  he reverses this construction.

  Kantor then transitions to the study of projective planes,
  in order to obtain an entirely different type of geometric
  view of symplectic spreads. First he considers the
  definition of the a spread again. Let $V'$ be a
  $2m$-dimensional vector space over $\GF(q)$. A
  \textit{spread} of $V'$ is a family $\Sigma'$ of $q^{m}+1$ 
  subspaces of dimension $m$ whose union is all of $V'$.
  This means that every nonzero vector is in a unique member
  of $\Sigma'$. Any family of $q^{m}+1$ $m$-spaces is a
  $2m$-space, any two of which have only $0$ in common, is a
  spread.

  Next he introduces the concept of an \textit{affine
  plane}. He states that the importance of spreads is that
  they produce affine planes. Let $A(\Sigma')$ denote the
  point-line geometry whose points are vectors and whose
  lines are the cosets $W + v$ with $W \in \Sigma'$, $v \in
  V'$. Then $A(\Sigma')$ is an affine plane of order
  $q^{m}$:
  \begin{itemize}
    \item Any two different points $u,v$ are on a unique
      line. (Namely, the line $W+v$ where $u-v \in W \in
      \Sigma'$).
    \item Given a line $L$ and a point $v$ not on it, there
      is a unique line through $v$ disjoint from $L$ 
      (namely, $W+v$ if $L$ is a coset of $W \in \Sigma'$).
    \item Each line has exactly $q^{m}$ points.
  \end{itemize}

  There is an obvious notion of parallelism, and by
  adjoingin a new line at infinity that contains all
  parallel classes we obtain a projective plane of order
  $q^{m}$. For each $c \in V'$ the translation $v \mapsto v
  + c$ is an automorphism fixing every parallel class. These
  affine planes are called translation planes.

  \begin{example}
    If $V'$ is a 2-dimensional vector space over $\GF(q)$,
    its set $\Sigma'$ of $1$-spaces is a desarguesian spread
    and $A(\Sigma')$ is called a desarguesian plane.
  \end{example}

  Desarguesian planes produce the original Kerdock codes. He
  adds the additional notion of a \textit{spread set}. Write
  $V' = X' \oplus Y'$ with $X',Y' \in \Sigma'$, and write
  matrices with respect to a basis of $V'$ consisting of a
  basis of $X'$ and basis of $Y'$. Then every member of
  $\Sigma' - \{Y\}$ can be written uniquely in the form $X'
  \begin{pmatrix} I & P \\ O & I \end{pmatrix} = ``y = xP''$
  for an $m \times m$ matrix $P$. The set $\mathcal S$ of
  such matrices is essentially what is called a
  \textit{spread set}, and they describe planes in a very
  familiar manner using the lines:
  \[
    ``x = c''
    \quad
    \text{and}
    \quad
    ``y = xP + b'',
    \quad b,c, \in V', P \in \mathcal S.
  \] 
  The spread $\Sigma'$ and the spread set $\mathcal S$ 
  determine one another in the obvious manner. In an
  analogous manner, he now connects binary operations and
  spreads. A binary operation $*$ satisfying the conditions
  of left distributivity and non divisibility of zero also
  determines a spread, consisting of $Y'$ and the subsets
  ``$y=x * s$'' of  $\GF(2^{m}) \oplus \GF(2^{m})$. The
  lines of the associated affine plane have a familiar
  appearance:
  \[
    ``x=c''
    \quad
    \text{and}
    \quad
    ``y=x*s+b'',
    \quad
    b,c,s \in \GF(2^{m}).
  \] 
  One can obtain a spread set defining $\mathcal S$ as the
  set of matrices of the maps $x \mapsto x * s$. Conversely,
  starting with a spread $\Sigma'$ in a vector space of
  characteristic 2, and distinct $X',Y' \in \Sigma'$, choose
  any bases for $X'$ and $Y'$, and obtain a spread set
  $\mathcal S$ of $m \times m$ matrices as above. Fix an
  arbitrary bijection $s \mapsto P_s$ of $\GF(2^{m}) \to
  \mathcal S$ with $P_0 = 0$, and define $x * s := xP_s$ for
  all $x,y \in \GF(2^{m})$. The conditions of the binary
  operation are fulfilled.

  So there is a natural way to go back and forth between
  spreads, spread sets, translation planes and binary
  operations. It's now time to see how it works for
  symplectic spreads $\Sigma'$ and the corresponding planes
  $A(\Sigma')$ called \textit{symplectic translation
  planes}. First he connects the notions of symplectic
  spreads and symmetric spread sets. Starting from a
  symplectic spread $\Sigma'$, fix $X',Y' \in \Sigma'$, and
  a basis of $X'$, but this time choose the basis $Y'$ to be
  the dual basis. Then the resulting spread set $\mathcal S$ 
  consists of symmetric matrices. Conversely each spread set
  consisting of symmetric matrices produces a symplectic
  spread.

  The same can be done by relating binary operations with
  Kerdock sets. Starting with an adequate binary oepration
  $*$, the transformations $x \mapsto x * s$ are
  self-adjoint with respect to the inner product $T(xy)$ on
  $\GF(2^{m})$. In terms of an orthonormal basis of
  $\GF(2^{m})$ this means that $x \mapsto x * s$ is
  represented by a symmetric matrix $P_s$. And so one
  obtains a spread set consisting of symmetric matrices,
  which produce a symplectic spread.

  \begin{align*}
    \text{orthogonal spread}
    &\leftrightarrow
    \text{symplectic spread} \\
    \text{orthogonal spread}
    &\leftrightarrow
    \text{Kerdock set} \\
    \text{Kerdock set}
    &\leftrightarrow
    \text{symmetric spread set} \\
    \text{Kerdock set}
    &\leftrightarrow
    \text{binary operation}.
  \end{align*}
  Each object on the left is associated with many on the
  right. Each one on the right is associated essentially
  with just one on the left, under whatever notion of
  equivalence is appropriate.

  As a follow up section, Kantor starts talking about bounds
  for \textit{line-sets} in $\R^{N}$ with prescribed angles.
  It simple to embed $\Z_2^{N}$ into $\R^{N}$, induced by
  the obvious isomorphism $\Z_2^{N} \to \{\pm 1\}^{N}$.
  Therefore a code $C$ of length $N$ produces an example of
  a set of unit vectors; and extremal properties of sets of
  unit vectors imply ones for codes.

  If we start with any Kerdock set of $\mathcal K$ of
  matrices, let the corresponding code be  $C(\mathcal K)$ 
  and set $N = 2^{n}$. Then we can form the following unit
  vectors in $\R^{N}$, where coordinates are again indexed
  by vectors of $\Z_2^{n}$ and the exponents again are just
  Kerdock codewords:
  \begin{equation}
    \frac{1}{2^{n / 2}} 
    \left(
      (-1)^{Q_B(v)+s \cdot v + \varepsilon}
    \right)_{v \in \Z_2^{n}},
    \quad
    \text{where }
    B \in \mathcal K, s \in \Z_2^{n}, \varepsilon \in \Z_2.
  \end{equation}
  To this set we add the $N = 2^{n}$ standard basis vectors
  and their negatives, given a total of $N^2+2N$ vectors in
  $\R^{N}$. A different view is given as sets of lines in
  $\R^{N}$, these are just the $1$-spaces spanned by the
  vectors in the above list. The distances between codewords
  in the kerdock code imply that any two of these lines are
  either perpendicular or set at an angle of $\cos^{-1} 1 /
  \sqrt{N}$.

  Finally he gets to $\Z_4$-codes. A $\Z_4$-code of length
  $N$ is just a subset $C_4$ of $\Z_4^{N}$. It is
  \textit{linear} if it is an additive subgroup. While one
  could use the Hamming metric here, an important discovery
  was the the Lee metric leads to better results. Next one
  defines the \textit{Gray map}, which allows for the
  encoding of the elements of $\Z_4$ (or of the points $1,
  i, -1, -i$ in the complex plane) as pairs of
  \textit{binary digits}. The $2$-adic expansion $c = c_1 +
  2c_2$ of $c \in \Z_4$ defines maps $c \mapsto c_1$ and $c
  \mapsto c_2$ from $\Z_4$ to $\Z_2$. Extend these in an
  obvious way to maps $v \mapsto v_i$ from $\Z_4^{N}$ to
  $\Z_2^{N}$. Then binary codes are obtained from
  $\Z_4$-codes as their images under the Gray map $\varphi :
  \Z_4^{N} \to \Z_2^{2N}$ given by
  \[
    v \varphi = (v_2, v_1+v_2),
    \quad v \in \Z_4^{N}.
  \] 
  The Gray map is an isometry with respect to the Lee metric
  in $\Z_4^{N}$ and the Hamming metric in $\Z_2^{2N}$. There
  is more than one Gray map, one can be followed by a
  permutation of the coordinates of $\Z_2^{2N}$ in order to
  obtain another one.

  Kantor goes on to show how to construct $Z_4$-Kerdock
  codes using a Kerdockt set and the Gray map. He then notes
  that although $C(\mathcal K)$ is never linear, one of the
  most striking discoveries was that the $\Z_4$-version
  $C_4(\mathcal S_{\mathcal K})$ of $C(\mathcal K)$ can be
  linear. 

  \begin{theorem}
    If $\mathcal K$ is a Kerdock set arising as in
    Proposition 2.5 by means of a binary operation $*$, then
    $C_4(\mathcal S_{\mathcal K})$ is a linear $Z_4$-code if
    and only if $*$ is 2-sided distributive.
  \end{theorem}

  Kerdock and other $Z_4$-linear codes have the striking
  properites of being optimal from a combinatorial point of
  view and yet having linear descriptions that simplify both
  their study and implementation.

  Finally he talks about bounds for line-sets in $\C^{N'}$ 
  with prescribed angles. As in the case for the usual
  Kerdock codes, the $Z_4$-Kerdock codes produce line-sets
  in $\C^{N'}$ via the isomorphism $\Z_4^{N'} \cong \langle
  i\rangle^{N'}$, where $N' = 2^{m}$. Thus, the
  $Z_4$-Kerdock code produces the following set of lines of
  $\C^{N'}$:
  \begin{equation}
    \langle \left( 
      i^{F_P(v) + 2 \hat s \cdot \hat v + \varepsilon}
    \right)_{v \in \Z_2^{m}} \rangle,
    \quad
    \text{where }
    P \in \mathcal S_{\mathcal K}, s \in \Z_2^{m},
    \varepsilon \in \Z_4,
  \end{equation}
  along with the $1$-spaces spanned by the $N' = 2^{m}$ 
  standard basis vectors. These lines fall into $N'+1$ 
  orthonormal frames such that the angle between members of
  different frames is always $\cos^{-1} 1 / \sqrt{N}$.

  \section{$\Z_4$-Kerdock codes, symplectic spreads and
  complex lines-sets with prescribed angles}

  This is a summary of chapter 5 and forward of the CCKS
  paper. To start out we recall the definition of a
  symplectic spread.

  \begin{definition}
    A symplectic spread of $\overline{F}$ is a family
    $\Sigma'$ of $2^{m}+1$ totally isotropic $m$-spaces such
    that every point in $\overline{F}$ lies in a unique
    member of $\Sigma'$. 
  \end{definition}

  A symplectic spread is a spread in the conventional sense,
  where by conventional he cites the book by Dembowski. In
  that book in chapter 5, Dembowski states that the known
  construction techniques for finite planes all use a finite
  vector space in a more or less obvious, but always
  essential way. This is the reason that these constructions
  always lead to planes of prime-power order. It is ones of
  the major unsolved problems of the theory of finite
  geometry whether or not there also exist planes of
  non-prime-power order.

  A symplectic spread is a family $\Sigma'$ of
  $\sqrt{|W|}+1$ subspaces of size $\sqrt{|W|}$ of finite
  vector space $W$ such that every non-zero vector lies in
  exactly one member of $\Sigma'$. A spread $\Sigma'$ 
  determines an \textit{affine plane} $A(\Sigma')$ of order
  $\sqrt{|W|}$ as follows: points are vectors, and lines are
  cosets $A+w$ where $A \in \Sigma'$ and $w \in W$.

  The magic happens here. After replacing $\Sigma'$ by
  $\Sigma'^{\ell}$ for some $\ell \in L^{\sharp}$, we may
  assume $\overline{X}(V)$ and $\overline{Y}(V)$ are in
  $\Sigma'$. Any totally isotropic $m$-space $\overline{A}
  \in \Sigma'$, with $\overline{A} \neq \overline{Y}(V)$ can
  be written in the form $\overline{X}(V) \begin{pmatrix} I
  & P_A \\ O & I\end{pmatrix} =
  \{\overline{X(a)}\overline{Y}(aP_A) : a \in V\}$ for a
  unique binary symmetric $m \times m$ matrix $P_A$. Such a
  matrix induces an isometry of $\overline{F}$. In terms of
  the affine plane $A(\Sigma')$, the subspace
  $\{\overline{X}(a)\overline{Y}(aP_A) : a \in V\}$ can be
  though of as the line ``$y = xP_A$''. From here we can
  obtain
   \begin{equation}
    \mathcal S(\Sigma')
    := \{P_A : \overline{A} \in \Sigma' \setminus
    \{\overline{Y}(V)\}\},
  \end{equation}
  is a set of $2^{m}$ symmetric $m \times m$ matrices such
  that the difference of any two is non-singular. If we
  remove the requirement of symmetry, then such a condition
  is a definition of a spread set. The corresponding affine
  plane $A(\mathcal S(\Sigma'))$ can be viewed as having $V
  \oplus V$ as its set of points, while the lines are the
  sets of points $(x,y) \in V \oplus V$ having the familiar
  appearnce
  \[
    x = b, \quad
    y = xP_A + b
    \quad
    \text{with }
    b \in V 
    \text{ and }
    \overline{A} \in \Sigma' \setminus \{\overline{Y}(V)\}.
  \] 
  Now he considers the unitary geometry of the $2$-group
  $F$. These are just counterparts of the orthogonal spreads
  lemmas.

  \begin{lemma}
    \begin{itemize}
      \item The set $\{\langle e_v \rangle : v \in V\}$ is
        the set of irreducible submodules for $Y(V)$.
      \item The set $\{\langle e_b^{*} \rangle : b \in V\}$ 
        is the set of irreducible submodules for $X(V)$.
    \end{itemize}
  \end{lemma}

  \begin{lemma}
    Let $A$ and $B$ be subgroups of $F$ such that
    $\overline{A}$ $\overline{B}$ are totally isotropic
    $m$-dimensional subspaces of the symplectic space
    $\overline{F}$.
    \begin{itemize}
      \item The set $\mathcal F(A)$ of $A$-irreducible
        subspaces of $\C^{2^{m}}$ is an orthogonal frame.
      \item Assume that $\overline{A} \cap \overline{B} =
        0$, and let $u_1$ and $u_2$ be unit vectors in
        different members of $\mathcal F(A) \cup \mathcal
        F(B)$. If $u_1$ and $u_2$ are both in $\mathcal
        F(A)$ or are both in $\mathcal F(B)$, then
        $(u_1,u_2) = 0$ ; otherwise, $|(u_1,u_2)| = 2^{-m /
        2}$.
      \item The set $\mathcal F(A)$ is left invariant by
        $F$.
    \end{itemize}
  \end{lemma}
  Again the proofs follow by using the group $L^{\sharp}$ to
  assume that $A = X(V)$ and $B = Y(V)$.

  \begin{theorem}
    Let $\Sigma'$ be any symplectic spread of the symplectic
    space $\overline{F}$. Then
    \[
      \mathcal F(\Sigma')
      := \cup_{\overline{A} \in \Sigma'}
      \mathcal F(A)
    \] 
    consists of $2^{m}(2^{m}+1)$ 1-spaces in $\C^{2^{m}}$ 
    such that if $u_1$ and $u_2$ are unit vectors in
    different members of $\mathcal F(\Sigma')$, then
    $|(u_1,u_2)| = 0$ or $2^{-m / 2}$.
  \end{theorem}

  Towards the end he gives us some examples. The usual is
  given by the desarguesian spread, which consists of the
  1-spaces of $\GF(2^{m})$. Then he goes to the construction
  of the $\Z_4$-Kerdock code $\mathcal K_4(\Sigma')$, the
  one that produces the $\Z_4$-lienar code that appears in a
  paper by Hammons et al. Such construction requires the
  Galois ring $\GR(4^{m})$, an extension of $\Z_4$ of degree
  $m$, contaning a $(2^{m}-1)$-th root of unity.

  The authors then mention the work of Kantor on the
  construction of inequivalent orthogonal spreads for $m$ 
  odd and composite. Let $n > 1$ be odd, let $L =
  \GF(2^{rn})$, let $K = \GF(2^{r})$, and let $\tr: L \to K$ 
  be the trace map. For $a \in L$ let $P_a$ be the
  $K$-linear map $L \to L$ given by
  \[
    xP_a = a^2 x + a \tr(x) + \tr(ax).
  \] 
  Then $a \mapsto P_a$ is additive and if $a \neq 0$, $P_a$ 
  is non-singular. 

  At the end the authors study the case of extraspecial
  $p$-groups, symplectic spreads and complex line-sets with
  prescribed angles (section 11). Basically the results are
  all analogues of the previous work, except they are now
  working with $\C^{q}$. 
\end{document}
